Windowed least-squares model reduction for dynamical systems

### 2019-10-16 section4/spacetime_numerical.tex
latest

Leveraging the definition of the fully discrete objective function~\eqref{eq:obj_lms}, and assuming $N \nstepsArg{n} \ge \stdimArg{n}$, \methodAcronym\ leveraging a \spaceTimeAcronym\ basis sequentially computes the generalized coordinates $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$, that satisfy
 \begin{equation}\label{eq:obj_gen_lms_final_st}
\begin{split}
& \underset{\stgenstatey}{\text{minimize } }
\objectiveArgLMS{n} (\stbasisArg{n}{\timeWindowArg{n}{0}} \stgenstatey + \stateInterceptSTArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptSTArg{n}).%  \\
%&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n(\stateFOMArgnt{1})(0) &
%n=1. \end{cases}
\end{split} 
\end{equation}
% \begin{equation}\label{eq:obj_gen_lms_final_st}
%\begin{split}
%&\stgenstateArg{n} = \underset{\stgenstatey}{\text{arg\,min } }
%\objectiveArgLMS{n} (\stbasisArg{n}{\timeWindowArg{n}{0}} \stgenstatey + \stateInterceptArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptArg{n})  \\
%&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n(\stateFOMArgnt{1})(0) &
%n=1. \end{cases}\end{split} 
%\end{equation}
It is noted that the boundary conditions are automatically satisfied through the definition of the \spaceTimeAcronym\ trial subspace. The minimization problem~\eqref{eq:obj_gen_lms_final_st} can be solved, e.g, via the Gauss--Newton method.

\begin{remark}
For the limiting case where the window size spans the entire domain, $\DeltaSlabArg{1}\equiv T$, and uniform quadrature weights are used, and the weighting matrix $\stweightingMatArg{n} = \mathbf{I}$, direct \methodAcronym\ with a \spaceTimeAcronym\ trial subspace recovers ST-LSPG with $\lspgWeightingSTArg{\cdot} = \mathbf{I}$. For equivalence conditions in the case where $\lspgWeightingSTArg{\cdot} \ne \mathbf{I}$, the objective function~\eqref{eq:obj_gen_lms_final_st} can be modified to enable weighted space--time norm. 
\end{remark}

@@

discretization of the integral. We introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ \text{Rank}(\stweightingMatOneArg{n}))$ \EP{Originally this was divided by $N$, just realized this actually shouldn't be N if we do hyper-reduction; should be something like Rank($\mathbf{W}^n)$ Agree?} collocation points over the $n$th window 
$$\timeStartArg{n} \le \collocPointSTArg{n}{1} \le \cdots \le \collocPointSTArg{n}{\ncollocSTArg{n}} \le \timeEndArg{n}.$$

@@

The least-squares problem can again be solved, e.g., via the Gauss--Newton method. 

@@

space--time trial subspace yields a series of dense (non)linear systems to be solved over each window.

>>>>

\EP{I will rework this section depending on 4.2\\
Leveraging the definition of the fully discrete objective function~\eqref{eq:obj_lms}, and assuming $N \nstepsArg{n} \ge \stdimArg{n}$, \methodAcronym\ leveraging a \spaceTimeAcronym\ basis sequentially computes the generalized coordinates $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$, that satisfy
 \begin{equation}\label{eq:obj_gen_lms_final_st}
\begin{split}
& \underset{\stgenstatey}{\text{minimize } }
\objectiveArgLMS{n} (\stbasisArg{n}{\timeWindowArg{n}{0}} \stgenstatey + \stateInterceptSTArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptSTArg{n}).%  \\
%&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n(\stateFOMArgnt{1})(0) &
%n=1. \end{cases}
\end{split} 
\end{equation}
% \begin{equation}\label{eq:obj_gen_lms_final_st}
%\begin{split}
%&\stgenstateArg{n} = \underset{\stgenstatey}{\text{arg\,min } }
%\objectiveArgLMS{n} (\stbasisArg{n}{\timeWindowArg{n}{0}} \stgenstatey + \stateInterceptArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptArg{n})  \\
%&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n(\stateFOMArgnt{1})(0) &
%n=1. \end{cases}\end{split} 
%\end{equation}
It is noted that the boundary conditions are automatically satisfied through the definition of the \spaceTimeAcronym\ trial subspace. The minimization problem~\eqref{eq:obj_gen_lms_final_st} can be solved, e.g, via the Gauss--Newton method.

\begin{remark}
For the limiting case where the window size spans the entire domain, $\DeltaSlabArg{1}\equiv T$, and uniform quadrature weights are used, and the weighting matrix $\stweightingMatArg{n} = \mathbf{I}$, direct \methodAcronym\ with a \spaceTimeAcronym\ trial subspace recovers ST-LSPG with $\lspgWeightingSTArg{\cdot} = \mathbf{I}$. For equivalence conditions in the case where $\lspgWeightingSTArg{\cdot} \ne \mathbf{I}$, the objective function~\eqref{eq:obj_gen_lms_final_st} can be modified to enable weighted space--time norm. 
\end{remark}
}

@@

discretization of the integral. Towards this end we introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ \text{Rank}(\stweightingMatOneArg{n}))$ \EP{Originally this was divided by $N$, just realized this actually shouldn't be N if we do hyper-reduction; should be something like Rank($\mathbf{W}^n)$ Agree?} collocation points over the $n$th window, $ \{ \collocPointSTArg{n}{i} \}_{i=1}^{\ncollocSTArg{n}} \subseteq [\timeStartArg{n},\timeEndArg{n}]$, $n=1,\ldots,\nslabs$. 
%with $\timeStartArg{n} \le \collocPointSTArg{n}{1} \le \cdots \le \collocPointSTArg{n}{\ncollocSTArg{n}} \le \timeEndArg{n}.$

@@

@@

space--time trial subspace yields a series of dense systems to be solved over each window.

### 2019-10-16 section4/spatial_numerical.tex
latest

Direct approaches solve optimization problem~\eqref{eq:tclsrm} by
``transcribing" the infinite dimensional optimization problem into a finite
dimensional one by discretizing the state and objective functional in time.
The minimization problem is then reformulated as a (non)linear programming, or
optimization, problem. A variety of direct solution approaches exist, 
including collocation approaches, spectral  methods, and genetic algorithms.  
In the context of \spatialAcronym\ trial subspaces, the most straightforward direct solution approach consists of the 
following steps: (1) numerically discretize the FOM ODE (and hence the
\textit{integrand} of the objective function in problem~\eqref{eq:obj_gen_slab}) and 
(2) select a numerical quadrature rule to evaluate the \textit{integral} in~\eqref{eq:obj_gen_slab}.
To this end, we define time grids
$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}\subseteq[\timeStartArg{n},\timeEndArg{n}]$,
$n=1,\ldots,\nslabs$ that
satisfy 
$\timeStartArg{n}=\timeWindowArg{n}{0} \le \cdots \le \timeWindowArg{n}{\nstepsArg{n}} 
 = \timeEndArg{n}$ and, for notational purposes, we define $\timeWindowArg{n}{i} = \timeWindowArg{n-1}{\nstepsArg{n-1}+i}$ for $i < 0$ and $n=2,\ldots,\nslabs$.
%\KTC{use cdots, not
%hdots or ldots. Don't use $\defeq$ for the first and last times; these are
%just equalities, i.e., you happen to set them equal to those values, but they
%aren't identically the same:}
Figure~\ref{fig:slab_fig2} depicts such a discretization.
We now outline the direct solution approach for linear-multistep schemes; we note that the formulation for
other time-integration methods (e.g., Runge--Kutta) follows closely. 
\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.0cm 4.5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid_timesteps.pdf} 
	\caption{Depiction of the $\nstepsArg{n}+1$ time instances over each window. In the figure, $\nstepsArg{n} = 2$ for all $n$.} 
\label{fig:slab_fig2} 
\end{centering} 
\end{figure}
%We now outline several commonly employed temporal discretization 
%schemes, specifically linear multistep and collocation (Runge--Kutta) methods, that leverage this time-grid to discretize the FOM ODE and 
%objective functional.

>>>>

Direct approaches solve optimization problem~\eqref{eq:tclsrm} by
``transcribing" the infinite dimensional optimization problem into a finite
dimensional one by discretizing the state and objective functional in time.
The minimization problem is then reformulated as a (non)linear programming, or
optimization, problem. A variety of direct solution approaches exist, 
including collocation approaches, spectral  methods, and genetic algorithms.  
In the context of \spatialAcronym\ trial subspaces, the most straightforward direct solution approach consists of the 
following steps: (1) numerically discretize the FOM ODE (and hence the
\textit{integrand} of the objective function in problem~\eqref{eq:obj_gen_slab}) and 
(2) select a numerical quadrature rule to evaluate the \textit{integral} in~\eqref{eq:obj_gen_slab}.
To this end, we define time grids
$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}\subseteq[\timeStartArg{n},\timeEndArg{n}]$ \EP{Shouldn't this be subset? We can't recover the continuous time interval},
$n=1,\ldots,\nslabs$ that
satisfy 
$\timeStartArg{n}=\timeWindowArg{n}{0} \le \cdots \le \timeWindowArg{n}{\nstepsArg{n}} 
 = \timeEndArg{n}$ and, for notational purposes, we define $\timeWindowArg{n}{i} = \timeWindowArg{n-1}{\nstepsArg{n-1}+i}$ for $i < 0$ and $n=2,\ldots,\nslabs$.
%\KTC{use cdots, not
%hdots or ldots. Don't use $\defeq$ for the first and last times; these are
%just equalities, i.e., you happen to set them equal to those values, but they
%aren't identically the same:}
Figure~\ref{fig:slab_fig2} depicts such a discretization.
We now outline the direct solution approach for linear-multistep schemes; we note that the formulation for
other time-integration methods (e.g., Runge--Kutta) follows closely. 
\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.0cm 4.5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid_timesteps.pdf} 
	\caption{Depiction of the $\nstepsArg{n}+1$ time instances over each window. In the figure, $\nstepsArg{n} = 2$ for all $n$.} 
\label{fig:slab_fig2} 
\end{centering} 
\end{figure}
%We now outline several commonly employed temporal discretization 
%schemes, specifically linear multistep and collocation (Runge--Kutta) methods, that leverage this time-grid to discretize the FOM ODE and 
%objective functional.
